---
name: Zachary
surname: S.L. Foster
position: "PhD Graduate Student"
address: "Oregon State University, Corvallis Oregon"
phone: +1 503 559 9772
email: "zacharyfoster1989@gmail.com"
github: zachary-foster
linkedin: zachary-foster-251a5777
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::hyndman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(purrr)
library(dplyr)
```


# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where, 
  "Bachelor of Science in Microbiology with Minor in Chemistry", "2008-2012", "Oregon State University", "Corvallis, Oregon, United States",
  "Bachelor of Science in Botany with Minor in Chemistry", "2008-2012", "Oregon State University", "Corvallis, Oregon, United States", 
  "PhD in Molecular and Cellular Biology", "2013-Present", "Oregon State University", "Corvallis, Oregon, United States",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Employment

```{r}
library(tibble)
tribble(
  ~ Job, ~ Year, ~ Institution, ~ Where,
  "Bioinformatics Programmer", "2009-2012", "Oregon State University", "Corvallis, Oregon, United States",
  "Graduate Research Assistant", "2013-Present", "Oregon State University", "Corvallis, Oregon, United States",
) %>% 
  detailed_entries(Job, Year, Institution, Where)
```


# Teaching

I have lead and created the content for workshops at scientific meetings and taught the lab/recitation portion of classes.

### Classes

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2015, "Graduate teaching assistant for Cell and Molecular Biology (BI314)", "Oregon State University",
  2017, "Graduate teaching assistant for Introduction to Biology (BI211)", "Oregon State University",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

### Workshops

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2016, "Reproducible research in R ", "American Phytopathological Society annual meeting",
  2018, "Analysis of microbiome data in R", "International Congress of Plant Pathology",
  2018, "Analysis of microbiome data in R", "International Phytobiomes Conference",
  2019, "Analysis of microbiome data in R", "BioMove Research Training Group, Germany",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Awards

I have won awards for research results, talks, and posters.

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2016, "Second place in the VISTAS Ecology Visualization Contest", "Evergreen State College",
  2017, "Best lightning talk", "Spring CGRB conference, Oregon State University",
  2017, "Best poster", "OMBI Microbiome Research Forum, Oregon State University",
  2017, "Phytobiomes journal poster award", "American Phytopathological Society annual meeting",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Grants

I have been awarded grants for research and travel to scientific meetings.

```{r}
tribble(
  ~Year, ~Desc, ~ Amount,
  2016, "NSF Graduate Data Science Workshop travel grant", "$1,000",
  2016, "Anita Summers travel grant", "$1,000",
  2017, "Anita Summers travel grant", "$1,000",
  2018, "rOpenSci fellowship", "$50,000",
) %>% 
  brief_entries(when = Year, what = Desc, with = Amount)
``` 


# Selected presentations

I have given invited and submitted talks at domestic and international scientific conferences.  

### Invited

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2019, "Presented a talk on the analysis and biases of metabarcoding data", " BioMove Research Training Group, Germany",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

### Submitted

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2017, "Metacoder: an R package for parsing, visualization, and manipulation of taxonomic data", "Center for Genome Research and Biocomputing (CGRB) spring conference",
  2017, "The effects of cultivar, production system, and nursery on the composition of the rhizosphere microbiome of cultivated rhododendrons in Oregon", "American Phytopathological Society annual meeting",
  2018, "Taxa and metacoder: R packages for parsing, visualization, and manipulation of taxonomic data", "useR! conference",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Volunteering

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  "2009-2017", 'Taught K-12 Students during "Discovery Days" at nine events and created educational games', "Oregon State University",
  "2016", "Planted early blue violet at Nestucca Bay National Wildlife refuge for reintroduction of the endangered Oregon silverspot butterfly.", "Institute for Applied Ecology",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Skills

```{r}
tribble(
  ~Type, ~Desc,
  "Programming", "I am an expert in R programming and package development. I also have experience in Python, C++, PHP, and C#.", 
  "Web design", "I have created and maintained websites using Rmarkdown and Drupal.", 
  "2D/3D graphics", "I can effectively create and edit raster graphics (Gimp), vector graphics (Inkscape), and 3D models (Blender, Meshroom). I am also experienced in digital photography using DSLRs.", 
  "Reproducible research", "I can create fully scripted, self-contained, and version-controlled analysis workflows using Rmarkdown that generate websites containing all descriptions, code, and publication-ready graphics.", 
  "Version control", "I have extensive experience using Git and Github.", 
  "Bioinformatics", "I have extensive experience analyzing biological data using programming languages and shell-scripting on Linux high-performance computing clusters and personal computers.", 
  "Molecular biology", "I have experience extracting DNA from diverse sample types, quantifying DNA, running gels, conducting PCR, and culturing microorganisms.", 
  "Field research", "I have planned, led, and helped conduct field research in natural and agricultural settings.", 
) %>% 
  detailed_entries(when = Type, what = Desc)
```

# Products

```{r}
tribble(
  ~Type, ~Desc, ~Contribution,  ~ URL,
  "R package", 'Metacoder: manipulation and visualization of metabarcoding data', 'Creator and lead developer', "Github repository: https://github.com/grunwaldlab/metacoder.",
  "R package", 'Taxa: classes to store and manipulate taxonomic data in R', 'Cocreator and lead developer', "Github repository: https://github.com/ropensci/taxa.",
  "Workshop", 'Analysis of Microbiome Community Data in R', 'Creator', "Website: https://grunwaldlab.github.io/metacoder_documentation/workshop--00--introduction.html",
  "Workshop", 'Reproducible Science in R', 'Cocreator', "Website: http://grunwaldlab.github.io/Reproducible-science-in-R",
) %>% 
  detailed_entries(when = Type, what = Desc, with = Contribution, where = URL)
``` 

# Publications

```{r}
raw_refs <- RefManageR::ReadBib('bibliography.bib')
sorted_pubs_path <- file.path('intermediate', 'sorted_publications.bib')
tibble(bibtype = "Article", 
       author = map_chr(raw_refs$author, paste, collapse = ', '),
       title = as.character(raw_refs$title),
       journaltitle = as.character(raw_refs$journal), 
       year = as.integer(raw_refs$year)) %>%
  arrange(desc(year)) %>%
  mutate(key = row_number()) %>%
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib(sorted_pubs_path)
  
bibliography_entries(sorted_pubs_path)
```

