---
name: Zachary
surname: S.L. Foster
position: "Postdoctoral Researcher"
address: "810 NW 23rd St, Corvallis, Oregon"
phone: +1 503 559 9772
email: "zacharyfoster1989@gmail.com"
github: zachary-foster
linkedin: zachary-foster-251a5777
date: "`r format(Sys.time(), '%B %Y')`"
output: vitae::hyndman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(purrr)
library(dplyr)
```


# Education

```{r}
library(tibble)
tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where, 
  "Bachelor of Science in Microbiology with a Minor in Chemistry", "2008-2012", "Oregon State University", "Corvallis, Oregon, United States",
  "Bachelor of Science in Botany", "2008-2012", "Oregon State University", "Corvallis, Oregon, United States", 
  "PhD in Molecular and Cellular Biology", "2012-2020", "Oregon State University", "Corvallis, Oregon, United States",
) %>% 
  detailed_entries(Degree, Year, Institution, Where)
```

# Employment

```{r}
library(tibble)
tribble(
  ~ Job, ~ Year, ~ Institution, ~ Where,
  "Bioinformatics programmer", "2009-2012", "Oregon State University", "Corvallis, Oregon, United States",
  "Graduate research assistant", "2012-2020", "Oregon State University", "Corvallis, Oregon, United States",
  "USDA Postdoctoral Researcher", "2020-current", "USDA, ARS, Horticultural Crops Research Unit", "Corvallis, Oregon, United States",
) %>% 
  detailed_entries(Job, Year, Institution, Where)
```

# Skills

I have diverse abilities and enjoy learning new skills.

```{r}
tribble(
  ~Type, ~Desc,
  "Programming", "I am an expert in R, Python, and Nextflow programming and software development. I also have experience in Bash, JavaScript, HTML, CSS, Java, Groovy, Snakemake, C#, C++, Perl, and PHP.", 
  "Software development", "I have experience in industry standard techniques including version control (Git), parallel programming, unit testing, performance profiling, code coverage profiling, CI/CD workflows (GitHub Actions), and containerization (Docker, Singularity, Conda).",
  "Bioinformatics", "I have extensive experience analyzing next generation sequencing data using programming languages and shell-scripting on high-performance computing clusters and personal computers. I use Linux daily and am comfortable administering Linux servers.", 
  "Reproducible research", "I can create fully scripted, self-contained, and version-controlled analysis workflows that generate HTML/PDF reports containing descriptions, code, and publication-ready graphics.", 
  "Molecular biology", "I have experience with high-throughput sequencing, extracting DNA from diverse sample types, quantifying DNA, running gels, conducting PCR, and culturing microorganisms.", 
  "Communication", "I have published manuscripts in high-impact journals, written extensive online docmentation, and presented my work at many conferences to a diverse audience.",
  "Teaching/Mentoring", "I have experience teaching people of all ages a variety of subjects and have created physical and online educational resources. I have also mentored many undergraduate research assistants working under my supervision.",
  "Field research", "I have planned, led, and helped conduct field research in natural and agricultural settings.", 
  "Web design", "I have created and maintained websites and managed web servers.", 
  "2D/3D graphics", "I can effectively create and edit raster graphics (Gimp), vector graphics (Inkscape), and 3D models (Blender, Meshroom, Freecad). I am also experienced in advanced digital photography techniques, including photogrammetry, focus stacking, and the creation of HDRIs.", 
) %>% 
  detailed_entries(when = Type, what = Desc, with = NA_character_, where = NA_character_)
```

# Selected products

I create and maintain software and educational material.

```{r}
tribble(
  ~Type, ~Desc, ~Contribution,  ~ URL,
  "Pipeline", 'Pathogensurveillance: A population genomics pipeline for biosurveillance.', 'Creator and lead developer', "Documentation: https://nf-co.re/pathogensurveillance",
  "R package", 'Metacoder: manipulation and visualization of metabarcoding data', 'Creator and lead developer', "GitHub repository: https://github.com/grunwaldlab/metacoder",
  "R package", 'Taxa: classes to store and manipulate taxonomic data in R', 'Cocreator and lead developer', "GitHub repository: https://github.com/ropensci/taxa",
  "Python package", 'Krisp: An efficient tool for finding diagnostic regions in whole genome data.', 'Cocreator and lead developer', "GitHub repository: https://github.com/grunwaldlab/krisp",
  "JavaScript library", 'Heat-tree: An interactive and embeddable phylogenetic tree viewer/editor', 'Creator and lead developer', "GitHub repository: https://github.com/grunwaldlab/heat-tree",
  "Workshop", 'Analysis of Microbiome Community Data in R', 'Creator', "Website: https://grunwaldlab.GitHub.io/metacoder_documentation",
  "Workshop", 'Reproducible Science in R', 'Cocreator', "Website: https://grunwaldlab.GitHub.io/Reproducible-science-in-R",
) %>% 
  detailed_entries(when = Type, what = Desc, with = Contribution, where = URL)
``` 


# Teaching

I have lead and created the content for workshops given at scientific meetings and taught the lab portion of classes.

### Classes

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2015, "Graduate teaching assistant for Cell and Molecular Biology (BI314)", "Oregon State University",
  2017, "Graduate teaching assistant for Introduction to Biology (BI211)", "Oregon State University",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

### Workshops

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2016, "Reproducible research in R ", "American Phytopathological Society annual meeting",
  2018, "Analysis of microbiome data in R", "International Congress of Plant Pathology",
  2018, "Analysis of microbiome data in R", "International Phytobiomes Conference",
  2019, "Analysis of microbiome data in R", "BioMove Research Training Group, Germany",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Awards

I have won awards for research results, talks, and posters.

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2016, "Second place in the VISTAS Ecology Visualization Contest", "Evergreen State College",
  2017, "Best lightning talk", "Spring Center for Genome Research and Biocomputing conference, Oregon State University",
  2017, "Best poster", "Microbiome Research Initiative (OMBI) Forum, Oregon State University",
  2017, "Phytobiomes journal poster award", "American Phytopathological Society annual meeting",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Grants

I have been awarded grants for research and travel to scientific meetings.

```{r}
tribble(
  ~Year, ~Desc, ~ Amount,
  2016, "NSF Graduate Data Science Workshop travel grant", "$1,000",
  2016, "Anita Summers travel grant", "$1,000",
  2017, "Anita Summers travel grant", "$1,000",
  2018, "rOpenSci fellowship", "$50,000",
) %>% 
  brief_entries(when = Year, what = Desc, with = Amount)
``` 


# Selected presentations

I have given invited and submitted talks at domestic and international scientific conferences.  

### Invited

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2019, "Ran a workshop on the analysis and biases of metabarcoding data", " BioMove Research Training Group, Germany",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

### Submitted

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  2017, "Metacoder: an R package for parsing, visualization, and manipulation of taxonomic data", "Center for Genome Research and Biocomputing (CGRB) spring conference",
  2017, "The effects of cultivar, production system, and nursery on the composition of the rhizosphere microbiome of cultivated rhododendrons in Oregon", "American Phytopathological Society annual meeting",
  2018, "Taxa and metacoder: R packages for parsing, visualization, and manipulation of taxonomic data", "useR! conference",
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 

# Service

```{r}
tribble(
  ~Year, ~Desc, ~ Institution,
  "2009-2017", 'Taught K-12 Students during "Discovery Days" at nine events and created educational games', "Oregon State University",
  "2016", "Planted early blue violet at Nestucca Bay National Wildlife refuge for reintroduction of the endangered Oregon silverspot butterfly", "Institute for Applied Ecology",
  "2020", "Planted native shrubs and trees at Powell Butte Nature Park, Portland, Oregon", "The Nature Conservancy",
  "2022", "Planted camas at Champoeg State Heritage Area, Oregon", "Institute for Applied Ecology",
  "2023", "Worked to restore alpine meadows at Mary's Peak, Oregon", "Institute for Applied Ecology",
  "2024", "Monitered alpine meadow restoration progress at Mary's Peak, Oregon", "Institute for Applied Ecology and US Forest Service"
) %>% 
  brief_entries(when = Year, what = Desc, with = Institution)
``` 


# Publications

```{r}
raw_refs <- RefManageR::ReadBib('bibliography.bib')
sorted_pubs_path <- file.path('intermediate', 'sorted_publications.bib')
tibble(bibtype = "Article", 
       author = map_chr(raw_refs$author, paste, collapse = ', '),
       title = as.character(raw_refs$title),
       journaltitle = as.character(raw_refs$journal), 
       year = as.integer(raw_refs$year)) %>%
  arrange(desc(year)) %>%
  mutate(key = row_number()) %>%
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib(sorted_pubs_path)
  
bibliography_entries(sorted_pubs_path)
```

